---
title: "Reproducible Research Project 1"
author: "Lisette"
date: "2/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/Lisette/Data science/Reproducible research")
```

## Reproducible Research Project 1

The data for this project can be found here: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
The variables included in this dataset are:

-steps: Number of steps taking in a 5-minute interval (missing values are coded as ğ™½ğ™°)
-date: The date on which the measurement was taken in YYYY-MM-DD format
-interval: Identifier for the 5-minute interval in which measurement was taken
###librarys
```{r}
library(dtplyr)
library(dplyr)
library(ggplot2)
library(sqldf)
library(lubridate)
```

###read in dataset and process
```{r echo=TRUE}
activity <- read.csv("activity.csv")
dim(activity)
names(activity)
```
###Histogram of the total number of steps taken each day
```{r}
#need to sum the steps per day
steps_by_date<-summarise(group_by(activity, date), daily_steps=sum(steps))
hist(steps_by_date$daily_steps, xlab="Daily steps", main="Number of Steps Taken Each Day",breaks=seq(from=0, to=25000, by=2500))
```


##Mean and median number of steps taken each day
```{r}

summarise(filter(steps_by_date, daily_steps>=0), Mean_daily_steps=mean(daily_steps) )
summarise(filter(steps_by_date, daily_steps>=0), Median_daily_steps=median(daily_steps) )
```

##Time series plot of the average number of steps taken
```{r}
#first we have to find average steps at each interval
average_by_interval <- summarise(group_by(na.omit(activity),interval), average_steps=mean(steps))
ggplot(data=average_by_interval, aes( interval,average_steps)) + geom_line() +   xlab("Interval") + ylab("Average Steps")
```
##The 5-minute interval that, on average, contains the maximum number of steps
```{r}
filter(average_by_interval, average_steps==max(average_by_interval$average_steps))
```
##Code to describe and show a strategy for imputing missing data
```{r}
#how many missing values of steps do we have? 
sum(is.na(activity$steps))

```
```{r}
#impute missing value by using mean for that interval
#identify which rows are missing
missing<- is.na(activity$steps)

#calculate average steps at each interval
avg_interval <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE, simplify=TRUE)
#create new data set with imputed average steps by interval for each missing value of steps
activity_full<-activity
activity_full$steps[missing] <- avg_interval[as.character(activity$interval[missing])]

```

##Histogram of the total number of steps taken each day after missing values are imputed
```{r}
steps_by_date_full<-summarise(group_by(activity_full, date), daily_steps=sum(steps))
hist(steps_by_date_full$daily_steps, xlab="Daily steps", main="Number of Steps Taken Each Day",breaks=seq(from=0, to=25000, by=2500))

#mean and median of full dataset with imputed values
summarise(filter(steps_by_date_full, daily_steps>=0), Mean_daily_steps=mean(daily_steps) )
summarise(filter(steps_by_date_full, daily_steps>=0), Median_daily_steps=median(daily_steps) )
```
##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
#read date field in as a date
activity_full$date<-ymd(activity_full$date)
  
#identify which days are weekdays and which are weekends
activity_full1<-mutate(activity_full, weektype = ifelse(weekdays(activity_full$date) == "Saturday" | weekdays(activity_full$date) == "Sunday", "weekend", "weekday"))

#First group by weekdaytype and then get means then create plots
by_weekdaytype <- activity_full1 %>%  group_by(interval, weektype) %>%  summarise(steps = mean(steps))
ggplot(by_weekdaytype, aes(x=interval, y=steps, color = weektype)) +
  geom_line() + facet_wrap(~weektype, ncol = 1, nrow=2)

```
